---
title: "1/24/2024 Lecture"
output: html_notebook
---

```{r}
library(ISLR2)
library(tidyverse)
```

# Correlations

## Model Problem

Create data

$$
y_i = x_i^3 + \epsilon_i
$$

Think of $\epsilon$ as "noise" in the measurements, and for simplicity, assume $\epsilon_i \sim N(0, \sigma^2)$ for some particular variance, $\sigma$ .

```{r}
set.seed(100) # set a seed for reproducibility
x <- seq(-5, 5, length=101) # 101 uniformly spaced points between -5 and 5
xi <-  rnorm(length(x)) # create a vector of Gaussians with same size as x, N(0,1)
df
# look at different magnitude noises
sigma2_vals = c(0, 10, 50,100, 500) 

# store in a data structure, data frame, with the columns we want
df <- tibble(x = numeric(), y=numeric(), noise=factor())
df
```

```{r}
#populate the data frame
for(j in 1:length(sigma2_vals)){
  sigma2 = sigma2_vals[j] # get the jth coordinate
  y <- x^3 + (sigma2) * xi # add noise with variance sigma2 to x^3
  df <- add_row(df, x = x, y = y, noise = as.factor(sigma2))
}
df
```

```{r}
correlations=c() # empty to start
for(j in 1:length(sigma2_vals)){
  sigma2 = sigma2_vals[j] # look at just the j-th noise case
  correlations <- append(correlations, 
                         cor(filter(df, noise == sigma2)$x,
                             filter(df, noise == sigma2)$y))
}
correlations
```

```{r}
cubic.plt=ggplot(df) + geom_point(aes(x,y, color=noise))
cubic.plt = cubic.plt + scale_color_discrete(name="Noise Intensity")
print(cubic.plt)
```

## Auto Data Set Problem

```{r}
Auto$cylinders = as_factor(Auto$cylinders)
Auto$origin = as_factor(Auto$origin)
Auto$year = as_factor(Auto$year)
```

```{r}
cor(select_if(Auto, is.numeric))
```

```{r}
pairs(select_if(Auto, is.numeric))
```

# More on ggplot

## Pair Correlations

A `ggplot`, `ggpairs`, provides better pair correlation figures, but it requires the `GGally` package to be installed

```{r}
library(GGally)
```

```{r}
auto.pairs.plt = ggpairs(select_if(Auto, is.numeric)) + 
  ggtitle("Pair Correlations for Auto")
print(auto.pairs.plt)
```

## Facets and Grids

Layout subplots

```{r}
Auto.facet.gplt = ggplot(Auto) + geom_point(aes(horsepower, mpg))
Auto.facet.gplt = Auto.facet.gplt + facet_wrap(~cylinders, nrow=2)
Auto.facet.gplt = Auto.facet.gplt + ggtitle("Horsepower vs mpg, by Cylinders")
print(Auto.facet.gplt)
```

We can use `facet_grid` to split along two different categorical variables

```{r}
Auto.grid.gplt = ggplot(Auto) + geom_point(aes(horsepower, mpg))
Auto.grid.gplt = Auto.grid.gplt + facet_grid(origin~cylinders)
Auto.grid.gplt = Auto.grid.gplt+ ggtitle("Horsepower vs mpg, by Cylinders and Origin")
print(Auto.grid.gplt)
```

## QQ Plots (revisited)

## Annotations (revisited)
