---
title: "1/12/2024 Lecture"
output: html_notebook
---

# Elementary Statistics in R

Generate a basic numeric data set:

```{r}
set.seed(123)
xdata = rnorm(100)
print(xdata)
```

Basic summary stastistics of data:

```{r}
summary(xdata)
```

More specific quantities:

```{r}
print(mean(xdata))
```

```{r}
print(median(xdata))
```

```{r}
print(var(xdata))
```

```{r}
print(sd(xdata))
```

```{r}
print(quantile(xdata))
```

By default, `quantile` gives the 4 quartiles. Suppose we want the quintiles, instead, values at 0%, 20%, 40%, 60%, 80%, 100%:

1.  Create an array of values, `pvals` that contains the values at which we want the quantile function
2.  Evaludate `quantile(xdata, probs=pvals)`

```{r}
pvals = c(0, .2, .4, .6, .8, 1)
```

```{r}
quantile(xdata, probs = pvals)
```

```{r}
pvals2=c(0.2, .87, .99)
quantile(xdata, probs = pvals2)
```

# Control Structures in R

For loops, if-then-else's,

```{r}
for (j in 1:5){
  print(j)
}
```

# Intro to Data Frames (tibbles)

These are built in, but we're going to use `tibble`, part of the `tidyverse`

```{r}
library(tidyverse) # loads the entire package fore use, a la import from python
```

Create a data frame:

```{r}
cast=tibble("Names"=c("Jerry", "George", "Elaine", "Cosmo"), "Ages"=c(38, 38, 37, 999))
```

```{r}
cast
```

999 is a "garbage" value that really reflects that the measurement was unavailable. Try to clean out garbage values. Use `NA` value. Plays an analogous role to `Inf`

```{r}
cast$Names # access columns
```

```{r}
cast$Ages # access columns
```

```{r}
cast$Ages[4]=NA
```

```{r}
cast
```

```{r}
summary(cast)
```

This is an elementary data cleaning example.

# Read an Excel Sheet into R

This builds on a `tidyverse` function

```{r}
library(readxl) # this provides Excel reading functions
```

```{r}
grades = read_excel("../data/Grades.xlsx")
```

```{r}
grades
```

`<dbl>` refers to double precision floating point numbers (i.e., the reals).

```{r}
summary(grades)
```

We will return to how to get R to interpret different columns as different data types (i.e., categorical).

```{r}
grades[1,] # first row
```

```{r}
grades[15,] # 15th row
```

```{r}
grades[2] # with no comma, second column
```
