---
title: "Lecture 1/29/2024"
output: html_notebook
---

```{r}
library(tidyverse)
library(ISLR2)
```

# Gather and Spread

Make the sample stock data:

```{r}
set.seed(100)
stocks.df = tibble(date=as.Date('2024-01-01')+0:9, "Apple"=100+cumsum(rnorm(10,0,4)),
                   "Google"=25+cumsum(rnorm(10,0,2)),
                   "Wawa"=10+cumsum(rnorm(10,0,0.5)))
```

```{r}
stocks.df
```

Reasonable for tabular presentation of data.

Challenge to plotting:

```{r}
plt = ggplot(stocks.df,mapping=aes(date,Apple))+geom_point()
print(plt)
```

But it would be nice to get all stocks at once, color coded, etc.

## Gather

```{r}
stocks.long.df = stocks.df %>% gather(key=symbol, value=price,-date)
stocks.long.df
```

Main arguments here are:

-   `key` - What we will call the new column. Here, we went from the columns being named after the companies, to having a column that is the company name.

-   `value` - What we will call the new column that holds the values that were previously the entries in the columns above.

-   `-date` -Retain this feature for each entry, this is essential:

```{r}
stocks.df %>% gather(key=symbol, value=price)
```

By not doing the `-date` , this just stacks all the columns together.

```{r}
stocks.long.df$symbol=as_factor(stocks.long.df$symbol)
```

```{r}
plt = ggplot(stocks.long.df, aes(date, price, color=symbol))+geom_point()
print(plt)
```

## Spread

Suppose we want somethign more like the tabular format that we started with; again, the table looked better (easier to comprehend) than the `stocks.long.df` layout. `spread` undoes the `gather` :

```{r}
stocks.short.df = stocks.long.df %>% spread(key=symbol, value=price)
stocks.short.df
```

This makes a distinct feature (column) for each entry in the `symbol` column(feature).

# Kables

To make nicely formatted tables that can be repsented, we use `kables` . These can be exported to Word, HTML, and Latex.

```{r}
library(knitr) # this provies the kable
library(kableExtra) # this provides formatting features for kables
```

```{r}
kable(stocks.df)
```

```{r}
kable(Auto) # this prints row "names", 1,2 ,3, 4...
```

```{r}
kable(Auto, row.names = FALSE)
```

Formatting kables:

```{r}
stocks.kb = kable(stocks.df, caption="Stocks Example") %>% kable_classic("striped", font_size=14)
stocks.kb
```

```{r}
stocks.kb = kable(stocks.df, caption="Stocks Example") %>% kable_classic(full_width=FALSE,"striped", font_size=14)
stocks.kb
```

```{r}
stocks.kb = kable(stocks.df, caption="Stocks Example") %>% kable_material(full_width=FALSE,"striped", font_size=10)
stocks.kb
```
