---
title: "Lecture 1/22/2024"
output: html_notebook
---

# Q-Q Plots

```{r}
set.seed(100)
u = runif(10^3) # U(0,1)
v = runif(2000, min=10, max=25) # U(10,25)
w = runif(100) # U(0,1)
x = rnorm(10^4) # N(0,1)
y = rnorm(5000, mean=10, sd =2) #N(10,2^2)
```

Use `qqplot` to compare data sets:

```{r}
qqplot(u,v)
```

This is a line, but it is not $y=x$, so these are not the same distribution.

```{r}
qqplot(u,w)
```

This is close to $y=x$, and it should be

```{r}
qqplot(u,x)
```

Comparison of $N(0,1)$ with $U(0,1)$; this is really not a line now.

```{r}
qqplot(x,y)
```

Two different normals, they sit on a line, but it's not $y=x$

```{r}
qqnorm(x)
```

`qqnorm` checks a single data set against $N(0,1)$

# Adding additional annotations

```{r}
library(tidyverse)
library(ISLR2)
```

```{r}
Auto$origin=as_factor(Auto$origin)
Auto$cylinders=as_factor(Auto$cylinders)
```

```{r}
mpg.plt = ggplot(Auto) # we will reuse this 
mpg.plt1 = mpg.plt + geom_boxplot(mapping = aes(x=origin, y= mpg))
print(mpg.plt1)
```

```{r}
summary(Auto$origin)
```

This could be in a caption or the text, but can we get into the figure directly?

Add an annotation:

```{r}
mpg.plt2 = mpg.plt1 + annotate("text", x=c(1.1, 2.1, 3.1), y = c(30, 40, 45), 
                               label=sprintf("n=%d", summary(Auto$origin)),
                               color="red")
print(mpg.plt2)
```

What did we do:

-   `text` - says the figure annotation will be a text object

-   `x` and `y` - these are selected by guess and check as to where to put the labels

-   `label=sprintf` - this is C/C++ style string formatting.

Alternatively, turn the summary information into a data frame itself:

```{r}
summary.df= tally(group_by(Auto, origin))
summary.df
```

`tally` sums up the data in each case from the `group_by`.
