---
title: "Lecture 2/5/2024"
output: html_notebook
---

```{r}
library(tidyverse)
```

# More Tables

Exporting for Excel/Word/etc.

Standard approach is to turn into something that can exported as a CSV file, and then use `write_csv`

## Simple example

```{r}
set.seed(123)
xdata = rnorm(20)
xdata
```

Store as a data frame, first:

```{r}
xdata.df = tibble(x=xdata)
xdata.df
```

```{r}
write_csv(xdata.df, "simple1.csv")
```

This `.csv` can then be read into Excel/Word/etc.

### Getting the summary out

```{r}
summary(xdata.df)
```

This does NOT work

```{r}
summary.df = as_tibble(summary(xdata.df))
```

If we can't turn this into a data frame, we can't export as a `.csv`

Instead, we use the `summarize` command:

```{r}
xdata.df %>% summarise(mean(x))
```

```{r}
xdata.df %>% summarise(mean(x), median(x))
```

We can reconstruct the full `summary` output in tibble format:

```{r}
xdata.df %>% summarise(min(x), 
                       quantile(x,0.25),
                       median(x),
                       mean(x),
                       quantile(0,0.75),
                       max(x))
```

```{r}
xdata.summary.df = xdata.df %>% summarise(min(x), 
                       quantile(x,0.25),
                       median(x),
                       mean(x),
                       quantile(0,0.75),
                       max(x))
```

```{r}
write_csv(xdata.summary.df, "xsummary.csv")
```

Suppose we want transpose for exporting,

```{r}
t(xdata.summary.df)
```

```{r}
names(xdata.summary.df)
```

```{r}
cbind(stat=names(xdata.summary.df), t(xdata.summary.df))
```

```{r}
xdata.summary.df.t = as_tibble(t(xdata.summary.df))
```

```{r}
write_csv(xdata.summary.df.t, "xsummaryt.csv")
```

## mtcars example

```{r}
mtcars # built in data set
```

Convert to tibble, and keep the row names:

```{r}
mtcars.df = as_tibble(mtcars, rownames="model") # need the rownames to retain the information
mtcars.df
```

```{r}
summary(mtcars.df)
```

To the export the summary for each property, we first use `gather` :

```{r}
mtcars.long.df = gather(mtcars.df, key="property", value ="value", -model)
mtcars.long.df
```

This gets us closer to where we want.

```{r}
mtcars.long.df %>% group_by(property) %>% summarize(Mean=mean(value), 
                                                    Median=median(value))
```

```{r}
mtcars.stats = mtcars.long.df %>% 
  group_by(property) %>% 
  summarize(Mean=mean(value), 
            Median=median(value))
write_csv(mtcars.stats, "mtcars1.csv")
```

### Transposed layout

```{r}
mtcars.stats.t = as_tibble(t(mtcars.stats[-1]), rownames="stat")
mtcars.stats.t
```

But notice we've lost the labels on top.

```{r}
mtcars.stats$property
```

See the column names:

```{r}
names(mtcars.stats.t)
```

Exclude the first one, as it is correct:

```{r}
names(mtcars.stats.t[-1])
```

Reassign the names:

```{r}
names(mtcars.stats.t)[-1] = mtcars.stats$property
```

```{r}
mtcars.stats.t
```

This can now be exported
