---
title: "Lecture 1/10/2024"
output: html_notebook
---

# Generating Random Variables in R

## Uniform Random Variables

```{r}
runif(1)
```

```{r}
runif(10)
```

These examples are $U(0,1)$ examples. For $U(a,b)$, set the `min` and `max` values as desired:

```{r}
runif(1,min=-1, max=2) # U(-1,2) variable
```

```{r}
runif(5,min=-1, max=2) 
```

## Reproducibility and Seeds

```{r}
runif(1)
```

```{r}
runif(1)
```

Note that these are *pseudo*-random number generators. There is a deterministic algorithm lurking inside.

Reproducibility is achieved by setting the *seed*:

```{r}
set.seed(100)
```

```{r}
runif(5)
```

Reset the seed to the same value and call again:

```{r}
set.seed(100)
```

```{r}
runif(1)
```

```{r}
runif(2)
```

Changing the seed gives a difference sequence:

```{r}
set.seed(200)
```

```{r}
runif(1)
```

Set the seed at the beginning of any code using random number generation for reproducibility. Also essential for debugging and software development.

## Normal Random Variables

For $N(0,1)$:

```{r}
rnorm(1)
```

```{r}
rnorm(10)
```

In R, normal $N(m,\sigma^2)$ is parametrized by $m$ and $\sigma$, the standard dev:

```{r}
rnorm(5, 1, 0.1)
```

# Markdown 

Greek letters rendered in LaTeX:

$\sigma$, $\alpha$, $\beta$, $\gamma$ $\delta$, $\Sigma$, $\Xi$ , $\Psi$ $\psi$

For bits of code, use single quotes: `min` ,`f` , `mean`
