---
title: "1/19/2024 Lecture"
output: html_notebook
---

# Better Graphics with ggplot2

As noted, the default graphics package in R is ok, but not great. Substantially better, out of the box, graphics can be generated with the `ggplot2` , which is part of the `tidyverse` :

```{r}
library(tidyverse)
```

The way of thinking of `ggplot2` graphics is in terms of addition, in that you will add graphical elements together, with the `+` , incrementally improving your plot. Roughly, you should think about it in terms of

Figure = Data Set + Graphical Manipulations of the data + Annotations

While it is not essential to create a data set, as a data frame (tibble or otherwise), to use `ggplot2` , there are advantages to using data frames here too.

## Parabola Plot

First, we will plot the parabola, $x(1-x)$ , manually:

```{r}
x = seq(0,1,length=51)
y = x*(1-x)
xy.plt = ggplot(mapping = aes(x,y))
```

This has created a plot, `xy.plt` which has, as its data, the *aesthetic*, `aes` . At this point, we see nothing. We can get the figure on the screen with the `print` command:

```{r}
print(xy.plt)
```

We now have a figure, but nothing in it. This is because we have not told it how to, geometrically, interpret the data.

### Scatter Plot

We have several options:

```{r}
xy.plt.point = xy.plt+geom_point() # interpret the data with a scatter plot
print(xy.plt.point)
```

### Line Plot

```{r}
xy.plt.line = xy.plt+geom_line() # interpret the data with a line plot
print(xy.plt.line)
```

### Titles

We can add a title to the plot with `ggtitle` :

```{r}
xy.plt.title = xy.plt+geom_line()+ggtitle("Parabola Plot") # interpret the data with a line plot
print(xy.plt.title)
```

### Saving to Disk

Figures can be saved to disk with the `ggsave` command:

```{r}
ggsave("ggplt1.pdf", plot = xy.plt.title, width = 6, height = 4) # size in inches
```

### Working with Data Frames

As noted, `ggplot2` plays nicely with data frames. First, we will stick our data into a data frame and then build on that:

```{r}
xy.df=tibble(Position=x,Velocity=y) # These are sample variable names
xy.df
```

```{r}
df.plt = ggplot(xy.df) + geom_line(mapping=aes(x=Position,y=Velocity), color="darkorange1",lwd=2) + 
  ggtitle("Plotting with a data frame")
print(df.plt)
```

Notice that it automatically picks up the variable names from the data frame for the axes

## Visualizing the Auto Data Frame

The `Auto` data frame is part of the `ISLR2` package:

```{r}
library(ISLR2)
```

```{r}
Auto = as_tibble(Auto)
Auto
```

Before visualizing this, we will convert certain properties to categorical (factor) types:

```{r}
Auto$cylinders = as_factor(Auto$cylinders)
Auto$origin = as_factor(Auto$origin)
```

### Boxplots

This is a good data set to use boxplots on, as we can compare the fuel efficiency as a function of region of origin or number of cylinders in the engine:

```{r}
mpg.plt = ggplot(Auto) # we will reuse this 
mpg.plt1 = mpg.plt + geom_boxplot(mapping = aes(x=origin, y= mpg))
print(mpg.plt1)
```

```{r}
mpg.plt2 = mpg.plt + geom_boxplot(mapping = aes(x=cylinders, y= mpg))
print(mpg.plt1)
```

**NOTE** that we changed the horizontal grouping by changing the `x` argument to the `aes`. We can further slice our data, visualizing the breakdowns together:

```{r}
mpg.plt3 = mpg.plt + geom_boxplot(mapping = aes(x=cylinders, y= mpg, fill=origin))
print(mpg.plt3)

```

```{r}
mpg.plt4 = mpg.plt + geom_boxplot(mapping = aes(x=origin, y= mpg, fill=cylinders))
print(mpg.plt4)

```

What's interesting about this plot is that you see that certain geographic regions only produce cars with some cylinder counts; only 4 and 6 cylinder engines are produced in all three regions.

### Histograms

We can perform similar analysis with histograms

```{r}
hist.mpg.plt1 = mpg.plt + geom_histogram(aes(x=mpg), bins=10) + ggtitle("MPG Distribution")
print(hist.mpg.plt1)
```

**NOTE:** I manually set `bins=10` , as in 10 histogram bins, because a warning was generated with the default value

```{r}
hist.mpg.plt2 = mpg.plt + geom_histogram(aes(x=mpg,fill=origin)) + ggtitle("MPG Distribution by Origin")
print(hist.mpg.plt2)
```

Alternatively, we can set `color` :

```{r}
hist.mpg.plt3 = mpg.plt + geom_histogram(aes(x=mpg,color=origin), bins=10) + ggtitle("MPG Distribution by Origin")
print(hist.mpg.plt3)
```

And we can improve the readability by making the bars transparent (changing `alpha` ):

```{r}
hist.mpg.plt4 = mpg.plt + geom_histogram(aes(x=mpg,fill=origin), bins=10, alpha=0.7) + ggtitle("MPG Distribution by Origin")
print(hist.mpg.plt4)
```
