---
title: "1/17/2024 Lecture"
output: html_notebook
---

# Loading/Exploring Data Sets

```{r}
library(tidyverse)
library(readxl)
```

```{r}
grades=read_excel("../data/Grades.xlsx")
```

## Setting Data Types in Data Frames

```{r}
grades
```

Note that both columns are stored as `<dbl>` , this is the double precision data type from C/C++/Java/Python/MATLAB/etc.

```{r}
summary(grades)
```

**NOTE** the statistics on the Student (ID), are meaningless.

We will let R know that it is a categorical variable:

```{r}
grades$Student = factor(grades$Student)
```

The above converts the data type of the column to `<fctr>` type, which is for categorical variables.

```{r}
grades
```

We can also set the Grade type to be an integer:

```{r}
grades$Grade=as.integer(grades$Grade)
```

```{r}
grades
```

Now summarize:

```{r}
summary(grades)
```

## Repeat this with CSV type data

```{r}
grades.csv = read.csv("../data/Grades.csv")
```

```{r}
grades.csv
```

This has been read in as a `df`, the default R, data frame (**NOT** yet a tibble). Usually worth converting to a tibble:

```{r}
grades.csv=as.tibble(grades.csv)
```

```{r}
grades.csv
```

Fix the data type of Student:

```{r}
grades.csv$Student=as.factor(grades$Student)
```

`as.factor` and `as_factor` are the same thing. True of many commands in R, `as.integer` and `as_integer` .

## Exporting data

```{r}
x=seq(0,1,length=11) # create some data
```

```{r}
write.csv(x,file="../data/sample.csv")
```

Export summary data.

```{r}
grades.summary = summary(grades)
```

```{r}
write.csv(grades.summary,file="grades.summary.txt")
```

How to export to latex as a table

```{r}
library(xtable)
```

**NOTE** `xtable` expects a data frame type as its input argument. This can be a tibble or a default data frame. We must first map it to a data frame with `as.data.frame`

```{r}
xtable(as.data.frame(x))
```

Get into a text file:

```{r}
print.xtable(xtable(as.data.frame(x)), file="tex_table.txt")
```

# Plotting and Visualization

## Basic Plotting with Built in functions

Make some basic data

```{r}
x=seq(0,1,length=51)
y = x*(1-x) # this is a parabola, x vs. y
```

Plot the parabola

```{r}
plot(x,y)
```

**NOTE** The default plotting is too bland for anything but a quick visualization.

Add some color and annotations:

```{r}
plot(x,y, main="Example Plot", xlab="x", ylab="x(1-x)", type="l", col="blue", lwd=2)
```

The additional arguments:

-   `main` - plot title

-   `xlab` - x axis label

-   `ylab` - y axis label

-   `type` - line with `l`, instead. of a scatter plot (default)

-   `col` - color

-   `lwd` - linewidth in pts.

What are our color options?

```{r}
colors()
```

```{r}
plot(x,y, main="Example Plot", xlab="x", ylab="x(1-x)", type="l", col="mediumorchid1", lwd=2)
```

Suppose we want to just export the figure:

```{r}
pdf(file="plot1.pdf", height=5.5, width=4) # note, need to specify height/width (in inches)
plot(x,y, main="Example Plot", xlab="x", ylab="x(1-x)", type="l", col="mediumorchid1", lwd=2)
dev.off() # stop writing to the file with this command
```

If the pdf format is ever too big, use `png` instead.

## Boxplots

```{r}
set.seed(100)
x=rnorm(100, mean=10, sd=2) # generate some data
```

```{r}
boxplot(x)b
```

-   Solid line is the median

-   The box is the interquartile range, middle 50% of the data

-   Stems are the max/mins

-   If they show up, scatter points are "outliers"

Make this more interesting:

```{r}
boxplot(x, main="Sample Box Plot", horizontal = TRUE, 
        col="red", xlab="x", ylab="Samples 1", cex.lab=1.5, cex.axis=1.25, cex.main=2)
```

-   `horizontal` changes the orientation

-   `cex.lab` , `cex.axis` and `cex.main` - control the font sizes as percentages of the default.

## Histogram

```{r}
x=rnorm(10^4, mean=2, sd=2)
```

```{r}
hist(x)
```

Annotate and improve:

```{r}
hist(x, breaks=50, main="N(2,2^2) Data", xlab="x", col="snow2", prob=TRUE)
```

By setting `prob=TRUE` , we are estimating the density, as in the probability density, from the data. Compare against analytic solution:

```{r}
hist(x, breaks=50, main="N(2,2^2) Data", xlab="x", col="snow2", prob=TRUE)
curve(dnorm(x, mean=2, sd=2), add=TRUE, col="navy", lwd=2)
```

`curve` plots a curve, `dnorm` specifies the function is the density of the normal, withe specified `mean` and `sd` , `add=TRUE` adds it to the existing plot
